Evaluation of probabilistic photometric redshift estimation approaches for \textsc{LSST}

### 2019-12-17 paper/authors.tex
re-fix affiliations and modify Mahalanobis formulat in CMNN

\author[LSST Dark Energy Science Collaboration]{
\parbox{\textwidth}{
\Large
S.J.~Schmidt$^{1}$,
A.I.~Malz$^{2,3,4}$,
J.Y.H.~Soo$^{5,6}$,
I.A.~Almosallam$^{7,8}$,
M.~Brescia$^{9}$,
S.~Cavuoti$^{9,10}$,
J.~Cohen-Tanugi$^{11}$,
A.J.~Connolly$^{12}$,
J.~DeRose$^{13,14,15,16,17}$,
P.E.~Freeman$^{18}$,
M.L.~Graham$^{12}$,
K.G.~Iyer$^{19,20}$,
M.J.~Jarvis$^{21,22}$,
J.B.~Kalmbach$^{12}$,
E.~Kovacs$^{23}$,
A.B.~Lee$^{18}$,
G.~Longo$^{10}$,
C.B.~Morrison$^{12}$,
J.A.~Newman$^{24}$,
E.~Nourbakhsh$^{1}$,
E.~Nuss$^{11}$,
T.~Pospisil$^{18}$,
H.~Tranin$^{11}$,
R.H.~Wechsler$^{25,16,26}$,
R.~Zhou$^{15,24}$,
R.~Izbicki$^{27,28}$
\begin{center} (LSST Dark Energy Science Collaboration) \end{center}
}
\vspace{0.4cm}
\\
\parbox{\textwidth}{
%\scriptsize
$^{1}$ Department of Physics, University of California, One Shields Ave., Davis, CA, 95616, USA\\
$^{2}$ German Centre of Cosmological Lensing, Ruhr-Universitaet Bochum, Universitaetsstra{\ss}e 150, 44801 Bochum, Germany\\
$^{3}$ Center for Cosmology and Particle Physics, New York University, 726 Broadway, New York, 10003, USA\\
$^{4}$ Department of Physics, New York University, 726 Broadway, New York, 10003, USA\\
$^{5}$ School of Physics, University of Science Malaysia, 11800 USM, Pulau Pinang, Malaysia\\
$^{6}$ Department of Physics and Astronomy, University College London, Gower Street, London WC1E 6BT, UK\\
$^{7}$ King Abdulaziz City for Science and Technology, Riyadh 11442, Saudi Arabia\\
$^{8}$ Information Engineering, Parks Road, Oxford, OX1 3PJ, UK\\
$^{9}$ INAF-Astronomical Observatory of Capodimonte, Salita Moiariello 16, 80131 Napoli, Italy\\
$^{10}$ Department of Physics E. Pancini, University Federico II, via Cinthia 6, I-80126, Napoli, Italy\\
$^{11}$ Laboratoire Univers et Particules de Montpellier, Universit\'e de Montpellier, CNRS, Montpellier, France\\
$^{12}$ DIRAC Institute and Department of Astronomy, University of Washington, Box 351580, U.W., Seattle WA 98195, USA\\
$^{13}$ Santa Cruz Institute for Particle Physics, Santa Cruz, CA 95064, USA\\
$^{14}$ Berkeley Center for Cosmological Physics, Department of Physics, University of California, Berkeley CA 94720\\
$^{15}$ Lawrence Berkeley National Laboratory, 1 Cyclotron Road, Berkeley, CA 94720, USA\\
$^{16}$ Kavli Institute for Particle Astrophysics and Cosmology and Department of Physics, Stanford University, Stanford, CA 94305, USA\\
$^{17}$ Department of Particle Physics and Astrophysics, SLAC National Accelerator Laboratory, Stanford, CA 94305, USA\\
$^{18}$ Department of Statistics \& Data Science, Carnegie Mellon University, 5000 Forbes Avenue, Pittsburgh, PA 15213, USA\\
$^{19}$ Dunlap Institute for Astronomy \& Astrophysics, University of Toronto, 50 St. George Street, Toronto, ON M5S 3H4, Canada\\
$^{20}$ Department of Physics and Astronomy, Rutgers, The State University of New Jersey, 136 Frelinghuysen Road, Piscataway, NJ 08854-8019 USA\\
$^{21}$ Astrophysics, Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford, OX1 3RH, UK\\
$^{22}$ Department of Physics and Astronomy, University of the Western Cape, Bellville 7535, South Africa\\
$^{23}$ Argonne National Laboratory, Lemont, IL 60439, USA\\
$^{24}$ Department of Physics and Astronomy and the Pittsburgh Particle Physics, Astrophysics and Cosmology Center (PITT PACC), University of Pittsburgh, Pittsburgh, PA 15260, USA\\
$^{25}$ Department of Physics, Stanford University, 382 Via Pueblo Mall, Stanford, CA 94305, USA\\
$^{26}$ SLAC National Accelerator Laboratory, Menlo Park, CA, 94025, USA\\
$^{27}$ Department of Statistics, Federal University of Sao Carlos, Sao Carlos, Brazil\\
$^{28}$ External collaborator\\
}
}

>>>>

\author[LSST Dark Energy Science Collaboration]{
\parbox{\textwidth}{
\Large
S.J.~Schmidt$^{1}$,
A.I.~Malz$^{2,3,4}$,
J.Y.H.~Soo$^{5,6}$,
I.A.~Almosallam$^{7,8}$,
M.~Brescia$^{9}$,
S.~Cavuoti$^{9,10}$,
J.~Cohen-Tanugi$^{11}$,
A.J.~Connolly$^{12}$,
J.~DeRose$^{13,14,15,16,17}$,
P.E.~Freeman$^{18}$,
M.L.~Graham$^{12}$,
K.G.~Iyer$^{19,20}$,
M.J.~Jarvis$^{21,22}$,
J.B.~Kalmbach$^{12}$,
E.~Kovacs$^{23}$,
A.B.~Lee$^{18}$,
G.~Longo$^{10}$,
C.B.~Morrison$^{12}$,
J.A.~Newman$^{24}$,
E.~Nourbakhsh$^{1}$,
E.~Nuss$^{11}$,
T.~Pospisil$^{18}$,
H.~Tranin$^{11}$,
R.H.~Wechsler$^{25,16,26}$,
R.~Zhou$^{15,24}$,
R.~Izbicki$^{27,28}$,
and The LSST Dark Energy Science Collaboration
}
\vspace{0.4cm}
\\
\parbox{\textwidth}{
(Affiliations are listed at the end of the paper)
\vspace{0.8cm}
}
}

### 2019-12-17 paper/tex/methods.tex
remove superfluous comments from tex
re-fix affiliations and modify Mahalanobis formulat in CMNN

Another key difference is the treatment of non-detections in one or more bands: some codes ignore incomplete bands, while others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
There are varying conventions among machine learning-based codes for treatment of non-detections, and no one prescription dominates in the \pz\ literature.
However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the $u$-band, and thus should not dominate the code performance differences.
%The specific choices for each code affect the results and contribute to the implicit prior influencing their output.

%The authors welcome interest from those outside \lsstdesc\ to have their codes assessed in future investigations that build upon this one.

@@

We describe the algorithms and implementations of the model-based and data-driven codes in Sections~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively, with a straw-person approach included in Section~\ref{sec:trainz}.
% For each approach, we note how it accounts for photometric uncertainties, how it interprets negative fluxes, what its native output format of \pzpdf s is, and what, if any, preprocessing, including validation, of the prior information it performs.
% In the following sections, we use a consistent notation defined in Table~\ref{tab:variables}.
%
% \begin{table}
% 	\label{tab:variables}
% 	\caption{Definitions of variables used in outlining \pzpdf\ codes}
% 	\begin{tabular}{ll}
% 		\hline
% 		\bf Symbol & \bf Definition \\
% 		\hline
% 		\multirow{ 2}{*}{$b = 1,\dots,B$	& \multirow{ 2}{*}{number $B$ of photometric filters $b$;} \\
% 														& \lsst's $ugrizy$ filters used here correspond to $B=6$ \\
% 		\multirow{ 2}{*}{$m_{b} = 1,\dots,B$	& \multirow{ 2}{*}{number $B$ of photometric filters $b$;} \\
% 																										& \lsst's $ugrizy$ filters used here correspond to $B=6$ \\
% 		$z$											& redshift \\
% 		$T$											& galaxy SED \\
%
% 	\end{tabular}
% \end{table}

@@

Here we test \bpz-v 1.99.3 \citep{Benitez:00} with the DC1 template set of Section~\ref{sec:buzztemplates}.
To keep the number of free parameters manageable, the DC1 template set is pre-sorted by the rest-frame $u-g$ colour and split into three broad classes of SED template, equivalent to the E, Sp and Im/SB types.
The Bayesian prior term $p(T \vert m_{0})$ was derived directly from the DC1 training set, and the other term $p(z \vert T, m_{0})$ was chosen to be the best fit for the eleven free parameters from the functional form of \citet{Benitez:00}.
We use template interpolation, creating two linearly interpolated templates between each basis SED (sorted by rest-frame $u-g$ colour) by setting the parameter \texttt{INTERP=2}.
%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
Prior to running the code, the non-detection placeholder magnitude was replaced with an estimate of the 1-$\sigma$ detection limit for the undetected band as a proxy for a value close to the estimated sky noise threshold.

\subsubsection{EAZY}
\label{sec:eazy}
%(Rongpu Zhou)

@@

\subsubsection{LePhare}
\label{sec:lephare}
%(C\'ecile Roucelle, Eric Nuss, Johann Cohen-Tanugi)

Photometric Analysis for Redshift Estimate \citep[\lephare,][]{Arnouts:99,Ilbert:06} uses the $\chi^2$ of Equation~\ref{eq_temp_chi} to match observed colours with those predicted from a template set.
The template set can be semi-empirical or entirely synthetic.
The reported \pzpdf\ is an arbitrary normalization of the likelihood evaluated on the output redshift grid.
%\textsc{LePhare} has been used to produce the COSMOS2015 photo-$z$ catalogue \citep{Laigle:16}.

@@

% Some aspects of data treatment were left to the individual code runners, for example, whether and how to split the DC1 training set for validation.
% The codes considered treated
% Another key difference is the treatment of non-detections in one or more bands.
% Some codes choose to ignore a band, others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
% There are varying conventions among training-based codes for treatment of non-detections, and no one prescription dominates in the photo-$z$ literature.
% The specific choices for each code affect the results, and contribute to the implicit prior influencing their output.
% However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the u-band, and thus should not dominate the code performance differences.

\subsubsection{ANNz2}
\label{sec:annz2}
%(John Soo)

\annz \citep{Sadeh:16} supports several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree, and k-nearest neighbour (KNN) regression.
In addition to accounting for errors on the input photometry, \annz\ uses the KNN-uncertainty estimate of \citet{Oyaizu:08} to quantify uncertainty in the choice of method over multiple runs.
Using the Toolkit for Multivariate Data Analysis with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, \annz\ can return the results of running a single machine learning algorithm, a ``best'' choice of the results from simultaneously running multiple algorithms (based on evaluation the cumulative distribution functions of validation set objects), or a combination of the results of multiple algorithms weighted by their method uncertainties averaged over multiple runs.
%\textsc{ANNz2} is capable of producing both photo-$z$ point estimates and redshift posterior probability distributions $p(z)$.
% It can also perform classifications, and supports reweighting between samples.
% \annz\ propagates the intrinsic uncertainty on the input parameters and the uncertainty in the machine learning method to the expected photo-$z$ solution, averaged over multiple runs weighted based on the performance of each run.

@@

\subsubsection{Colour-Matched Nearest-Neighbours}
\label{sec:cmnn}
%(Melissa Graham)

The colour-matched nearest-neighbours photometric redshift estimator \citep[\cmnn,][]{Graham:17} uses a training set of galaxies with known redshifts that has equivalent or better photometry than the test set in terms of quality and filter coverage.
For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M = \sum_{j}^{N_{\mathrm{colours}}} (c^{\mathrm{train}}_{j} - c^{\mathrm{test}}_{j})^{2} / \delta c_{\mathrm{test}}^2$ in the space of available colours $c$, with colour measurement errors $\delta c_{\mathrm{test}}$ and $N_{\mathrm{colours}} = 5$ colours $j$ defined by the $ugrizy$ filters, which defines the set of colour-matched neighbours based on a value of the percent point function (PPF).
As an example, for $N_{\mathrm{filt}}=5$ with PPF$=0.95$, $95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
Undetected bands are dropped, thereby reducing the effective $N_{\mathrm{filt}}$ for that galaxy.
The \pzpdf\ of a given test set galaxy is the normalized distribution of redshifts of its colour-matched subset of training set galaxies.

@@

\subsubsection{FlexZBoost}
\label{sec:flexzboost}
%(Ann Lee, Rafael Izbicki, Taylor Pospisil, Peter Freeman)

@@

%\subsubsection{Frankenz}
%\label{sec:frankenz}
%(seeking volunteers)
%
%\red{\textsc{Frankenz}\footnote{\url{https://github.com/joshspeagle/frankenz}} \cite{Speagle:frankenz} is...}
%

\subsubsection{GPz}
\label{sec:gpz}
%(Ibrahim Almosallam)

\gpz \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process-based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, that produces input-dependent variance estimates corresponding to heteroscedastic noise.
The model assumes a Gaussian posterior probability $p(z \vert \x) = \mathcal{N}\left(z \vert \mu(\x), \sigma(\x)^{2}\right)$ of the output redshift $z$ given the input photometry $\x$.
The mean $\mu(\x)$ and the variance $\sigma(\x)^{2}$ are modeled as functions $f(\x) = \sum_{i=1}^{m}w_{i}\phi_{i}(\x)$ that are linear combinations of $m$ basis functions $\left\{\phi_{i}(\x)\right\}_{i=1}^{m}$ with associated weights $\left\{w_{i}\right\}_{i=1}^{m}$.
% Basis function models, for specific classes of basis functions such as the sigmoid or the squared exponential, have the advantage of being universal approximators, i.e. there exist a function of that form that can approximate any function, with mild assumptions, to any desired degree of accuracy.
The details on how to learn the parameters of the model and the hyper-parameters of the basis functions are described in \citet{Almosallam:15b}.
\gpz's variance estimate is composed of a model uncertainty term corresponding to sparsity of the training set photometry and a noise uncertainty term encompassing noisy photometric observations, enabling quantification of any need for more representative or more precise training samples.
\gpz\ may also weight training set samples by importance according to $|z_{\mathrm{spec}} - z_{\mathrm{phot}}| / (1+z_{\mathrm{spec}})$ to minimize the normalized \pz\ point estimate error.  However, this function may be adapted to \pzpdf s, adding weight to test set galaxies that are not well-represented in the training set.

To smooth the long tail in the distribution of magnitude errors, we use the logarithm of the magnitude errors, improving numerical stability and eliminating the need for constraints on the optimization process.
% We use principaphoto-\mathinhead{$z$}{z} interim posteriorsl component analysis (PCA) to decorrelate the data
% WHAT IS BEING DECORRELATED HERE?
Unobserved magnitudes $x_{\mathrm{u}} = \mu_{\mathrm{u}} + \Sigma_{\mathrm{uo}}\Sigma_{\mathrm{oo}}^{-1}(x_{\mathrm{o}} - \mu_{\mathrm{o}})$ were imputed from observed magnitudes $x_{\mathrm{o}}$ and the training set mean $\mu$ and covariance $\Sigma$ using a linear model.
This is the optimal expected value of the unobserved variables given the observed ones under the assumption that the distribution is jointly Gaussian; note that this reduces to a simple average if the covariates are independent with $\Sigma_{\mathrm{uo}} = 0$.
We reserved for validation 20 per cent of the training set and used the Variable Covariance option in \gpz\ with 200 basis functions (see \citet{Almosallam:15b} for details), and did not apply cost-sensitive learning options.

\subsubsection{METAPhoR}
\label{sec:metaphor}
%(Stefano Cavuoti, Massimo Brescia, Giuseppe Longo)

Machine-learning Estimation Tool for Accurate Photometric Redshifts \citep[\metaphor,][]{Cavuoti:17} is based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA) with the least square error model and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
%, already validated on photo-$z$'s in several cases \citep{de_Jong:17,Cavuoti:17b,Cavuoti:15,Brescia:14,Brescia:13,Biviano:13}.
\Pzpdf s are generated by running $N$ trainings on the same training set, or $M$ trainings on $M$ different random samplings of the training set.
Upon regression of the test set, the photometry $m_{ij}$ of each test set galaxy $j$ in filter $i$ is perturbed according to $m_{ij}' = m_{ij} + \alpha_{i} F_{ij} \epsilon$ in terms of the standard normal random variable $\epsilon \sim \mathcal{N}(0, 1)$, a multiplicative constant $\alpha_{i}$ permitting accommodation of multi-survey photometry, and a bimodal function $F_{ij}$ composed of a polynomial fit of the mean magnitude errors on the binned bands plus a constant term representing the threshold below which the polynomial's noise contribution is negligible \citep{Brescia:18}.
% At a higher level, the pipeline mainly consists of three modules: (i) \textit{data pre-processing}, including a catalogue cross-matching sub-module \citep[based on the tool C3, ][]{Riccio:17}, a sub-module for photometric evaluation and error estimation of the multi-band catalogue used as Knowledge Base (KB), and a sub-module dedicated to the perturbation of the photometric KB, propaedeutic to the PDF estimation; (ii) \textit{photo-$z$ prediction}, which is the training/validation/test phase, producing the photo-$z$'s point estimates, based on a pre-selected ML method; (iii) \textit{PDF estimation}, specifically designed to calculate the PDF of the photo-$z$ estimation errors.
% The last module includes also a post-processing tool, providing some statistics on the produced point estimates and PDFs.

@@

\subsubsection{SkyNet}
\label{sec:skynet}
%(J. Cohen-Tanugi and Hugo Tranin)

\skynet \citep{Graff:14} employs a neural network based on a second-order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}. %It has been used efficiently for redshift PDF estimates \citep{Sanchez:14,Bonnett:15,Bonnett:16}.
The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with 12 nodes corresponding to the 6 photometric magnitudes and their measurement errors.
% We use \skynet\ as a regressor for \pz\ point estimation and as a classifier for \pzpdf\ estimation.

% The regressor used a standard \chisq\ error function with a single linear node as the output layer and 10 nodes with a $\tanh$ activation function for each hidden layer.
\skynet's classifier mode uses a cross-entropy error function with a 20:40:40 node (all rectified linear units) architecture for each hidden layer and an output layer of 200 nodes corresponding to 200 bins for the PDF, with a softmax activation function to enforce the normalization condition that the probabilities sum to unity.
While previous implementations of the code \citep[see Appendix C.3 of~][]{Sanchez:14,Bonnett:15} implement a sliding bin smoothing, no such procedure was used in this study.

We pre-whitened the data by pegging the magnitudes to (45,45,40,35,42,42) and errors to (20,20,10,5,15,15) for $ugrizy$ filters, respectively.
% NO CLUDE WHAT THIS MEANS
To avoid over-fitting, $30$ per cent of the training set was reserved for validation, and training was halted as soon as the error rate began to increase on the validation set.
The weights were randomly initialized based on normal sampling.
% WHAT WEIGHTS?

\subsubsection{TPZ}
\label{sec:tpz}
%(Erfan Nourbakhsh)

@@

Though \trainz\ is strongly vulnerable to a nonrepresenative training set, it should optimize performance metrics probing the ensemble properties of the galaxy sample, modulo Poisson error due to small sample size, as the training set and test set are drawn from the same underlying population.
We will demonstrate its performance under the metrics of Section~\ref{sec:metrics} and discuss it as an illustrative experimental control case in Section~\ref{sec:caution} to highlight the limitations of our evaluation criteria for \pzpdf s.
% any attempts to break up the sample into tomographic bins will fail, as every galaxy has an identical $p(z)$.

>>>>

Another key difference is the treatment of non-detections in one or more bands: some codes ignore incomplete bands, while others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
There are varying conventions among machine learning-based codes for treatment of non-detections, and no one prescription dominates in the \pz\ literature.
However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the $u$-band, and thus should not dominate the code performance differences.

@@

We describe the algorithms and implementations of the model-based and data-driven codes in Sections~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively, with a straw-person approach included in Section~\ref{sec:trainz}.

@@

Here we test \bpz-v 1.99.3 \citep{Benitez:00} with the DC1 template set of Section~\ref{sec:buzztemplates}.
To keep the number of free parameters manageable, the DC1 template set is pre-sorted by the rest-frame $u-g$ colour and split into three broad classes of SED template, equivalent to the E, Sp and Im/SB types.
The Bayesian prior term $p(T \vert m_{0})$ was derived directly from the DC1 training set, and the other term $p(z \vert T, m_{0})$ was chosen to be the best fit for the eleven free parameters from the functional form of \citet{Benitez:00}.
We use template interpolation, creating two linearly interpolated templates between each basis SED (sorted by rest-frame $u-g$ colour) by setting the parameter \texttt{INTERP=2}.
Prior to running the code, the non-detection placeholder magnitude was replaced with an estimate of the 1-$\sigma$ detection limit for the undetected band as a proxy for a value close to the estimated sky noise threshold.

\subsubsection{EAZY}
\label{sec:eazy}

@@

\subsubsection{LePhare}
\label{sec:lephare}

Photometric Analysis for Redshift Estimate \citep[\lephare,][]{Arnouts:99,Ilbert:06} uses the $\chi^2$ of Equation~\ref{eq_temp_chi} to match observed colours with those predicted from a template set.
The template set can be semi-empirical or entirely synthetic.
The reported \pzpdf\ is an arbitrary normalization of the likelihood evaluated on the output redshift grid.

@@

\subsubsection{ANNz2}
\label{sec:annz2}

\annz \citep{Sadeh:16} supports several machine learning algorithms, including artificial neural networks (ANN), boosted decision tree, and k-nearest neighbour (KNN) regression.
In addition to accounting for errors on the input photometry, \annz\ uses the KNN-uncertainty estimate of \citet{Oyaizu:08} to quantify uncertainty in the choice of method over multiple runs.
Using the Toolkit for Multivariate Data Analysis with ROOT\footnote{\url{http://tmva.sourceforge.net/}}, \annz\ can return the results of running a single machine learning algorithm, a ``best'' choice of the results from simultaneously running multiple algorithms (based on evaluation the cumulative distribution functions of validation set objects), or a combination of the results of multiple algorithms weighted by their method uncertainties averaged over multiple runs.

@@

\subsubsection{Colour-Matched Nearest-Neighbours}
\label{sec:cmnn}

The colour-matched nearest-neighbours photometric redshift estimator \citep[\cmnn,][]{Graham:17} uses a training set of galaxies with known redshifts that has equivalent or better photometry than the test set in terms of quality and filter coverage.
For each galaxy in the test set, \cmnn\ identifies a colour-matched subset of training galaxies using a threshold in the Mahalanobis distance $D_M = \sum_{j}^{N_{\mathrm{colours}}} (c_{\mathrm{j,train}} - c_{\mathrm{j,test}})^{2} / \delta c_{\mathrm{j,test}}^2$ in the space of available colours $c$, with colour measurement errors $\delta c_{\mathrm{test}}$ and $N_{\mathrm{colours}} = 5$ colours $j$ defined by the $ugrizy$ filters, which defines the set of colour-matched neighbours based on a value of the percent point function (PPF).
As an example, for $N_{\mathrm{filt}}=5$ with PPF$=0.95$, $95$ per cent of all training galaxies consistent with the test galaxy will have $D_M < 11.07$.
Undetected bands are dropped, thereby reducing the effective $N_{\mathrm{filt}}$ for that galaxy.
The \pzpdf\ of a given test set galaxy is the normalized distribution of redshifts of its colour-matched subset of training set galaxies.

@@

\subsubsection{FlexZBoost}
\label{sec:flexzboost}

@@

\subsubsection{GPz}
\label{sec:gpz}

\gpz \citep{Almosallam:16a,Almosallam:15b} is a sparse Gaussian process-based code, a scalable approximation of full Gaussian Processes \citep{Rasmussen:06}, that produces input-dependent variance estimates corresponding to heteroscedastic noise.
The model assumes a Gaussian posterior probability $p(z \vert \x) = \mathcal{N}\left(z \vert \mu(\x), \sigma(\x)^{2}\right)$ of the output redshift $z$ given the input photometry $\x$.
The mean $\mu(\x)$ and the variance $\sigma(\x)^{2}$ are modeled as functions $f(\x) = \sum_{i=1}^{m}w_{i}\phi_{i}(\x)$ that are linear combinations of $m$ basis functions $\left\{\phi_{i}(\x)\right\}_{i=1}^{m}$ with associated weights $\left\{w_{i}\right\}_{i=1}^{m}$.
The details on how to learn the parameters of the model and the hyper-parameters of the basis functions are described in \citet{Almosallam:15b}.
\gpz's variance estimate is composed of a model uncertainty term corresponding to sparsity of the training set photometry and a noise uncertainty term encompassing noisy photometric observations, enabling quantification of any need for more representative or more precise training samples.
\gpz\ may also weight training set samples by importance according to $|z_{\mathrm{spec}} - z_{\mathrm{phot}}| / (1+z_{\mathrm{spec}})$ to minimize the normalized \pz\ point estimate error.  However, this function may be adapted to \pzpdf s, adding weight to test set galaxies that are not well-represented in the training set.

To smooth the long tail in the distribution of magnitude errors, we use the logarithm of the magnitude errors, improving numerical stability and eliminating the need for constraints on the optimization process.
Unobserved magnitudes $x_{\mathrm{u}} = \mu_{\mathrm{u}} + \Sigma_{\mathrm{uo}}\Sigma_{\mathrm{oo}}^{-1}(x_{\mathrm{o}} - \mu_{\mathrm{o}})$ were imputed from observed magnitudes $x_{\mathrm{o}}$ and the training set mean $\mu$ and covariance $\Sigma$ using a linear model.
This is the optimal expected value of the unobserved variables given the observed ones under the assumption that the distribution is jointly Gaussian; note that this reduces to a simple average if the covariates are independent with $\Sigma_{\mathrm{uo}} = 0$.
We reserved for validation 20 per cent of the training set and used the Variable Covariance option in \gpz\ with 200 basis functions (see \citet{Almosallam:15b} for details), and did not apply cost-sensitive learning options.

\subsubsection{METAPhoR}
\label{sec:metaphor}

Machine-learning Estimation Tool for Accurate Photometric Redshifts \citep[\metaphor,][]{Cavuoti:17} is based on the Multi Layer Perceptron with Quasi Newton Algorithm (MLPQNA) with the least square error model and Tikhonov $L_{2}$-norm regularization \citep{Hofmann:18}.
\Pzpdf s are generated by running $N$ trainings on the same training set, or $M$ trainings on $M$ different random samplings of the training set.
Upon regression of the test set, the photometry $m_{ij}$ of each test set galaxy $j$ in filter $i$ is perturbed according to $m_{ij}' = m_{ij} + \alpha_{i} F_{ij} \epsilon$ in terms of the standard normal random variable $\epsilon \sim \mathcal{N}(0, 1)$, a multiplicative constant $\alpha_{i}$ permitting accommodation of multi-survey photometry, and a bimodal function $F_{ij}$ composed of a polynomial fit of the mean magnitude errors on the binned bands plus a constant term representing the threshold below which the polynomial's noise contribution is negligible \citep{Brescia:18}.

@@

\subsubsection{SkyNet}
\label{sec:skynet}

\skynet \citep{Graff:14} employs a neural network based on a second-order conjugate gradient optimization scheme \citep[see][for further details]{Graff:14}. 
The neural network is configured as a standard multilayer perceptron with three hidden layers and one input layer with 12 nodes corresponding to the 6 photometric magnitudes and their measurement errors.

\skynet's classifier mode uses a cross-entropy error function with a 20:40:40 node (all rectified linear units) architecture for each hidden layer and an output layer of 200 nodes corresponding to 200 bins for the PDF, with a softmax activation function to enforce the normalization condition that the probabilities sum to unity.
While previous implementations of the code \citep[see Appendix C.3 of~][]{Sanchez:14,Bonnett:15} implement a sliding bin smoothing, no such procedure was used in this study.

We pre-whitened the data by pegging the magnitudes to (45,45,40,35,42,42) and errors to (20,20,10,5,15,15) for $ugrizy$ filters, respectively.
To avoid over-fitting, $30$ per cent of the training set was reserved for validation, and training was halted as soon as the error rate began to increase on the validation set.
The weights were randomly initialized based on normal sampling.

\subsubsection{TPZ}
\label{sec:tpz}

@@

Though \trainz\ is strongly vulnerable to a nonrepresenative training set, it should optimize performance metrics probing the ensemble properties of the galaxy sample, modulo Poisson error due to small sample size, as the training set and test set are drawn from the same underlying population.
We will demonstrate its performance under the metrics of Section~\ref{sec:metrics} and discuss it as an illustrative experimental control case in Section~\ref{sec:caution} to highlight the limitations of our evaluation criteria for \pzpdf s.
